A Formal Model of Checked C

### 2022-12-30 sandp23/evaluation.tex

% \review{While I found the idea behind section V very interesting, the current version
%   of this section lacks some details that would help in better understanding (1) 
%   how the approach works, and (2) the overall scope of the approach. 
%   $\\$
%   For instance, the authors state that, following [19], they try to "exercise
%   interesting patterns" by adding "admissible but redundant typing rules" like
%   G-ASTR. There are a few points that are unclear here: (1) are these rules
%   discovered manually or automatically (starting from the Redex semantics)?, (2)
%   are there any guiding principles for coming up with rules that lead to
%   interesting cases?
%   $\\$
%   Later, the authors refer to "generation rules modified to be slightly more
%   permissive" to generate "a little" ill-typed terms. Again, are these rules
%   obtained automatically or defined manually? If the latter, did you follow any
%   methodology to derive such rules? Are these rules the same as the "admissible
%   but redundant typing rules" from above?}
% \liyi{Deena? Leo? }

\begin{itemize}

\item provide evidence that the \systemname compiler is efficient. Compare the compiler with respect to other work, like RLBox, also the previous \checkedc compiler.

\item provide user experience of \systemname. We restrict the use of checked pointers compared to previous checked-c compiler. Is the restriction arrangable. We can say that the tainted shells are auto-matically generated, so we have a mechanism for auto-generating tainted pointers if necessary.

\item if we have space, we can re-introduce random testing a little, saying that how it helps us to develop the compiler.

\item we can then talk about the possible bugs we find in the \checkedc compiler for function pointer or the RLBox bugs.

\end{itemize}

% 
%\leo{The following is extremely
%  weak. ``Most of them'', were there any that weren't? Which ones? For
%  the ones that were, mention github issues.}  The random generator,
%equipped with the conversion tool, successfully found a few minor
%errors in the clang compiler, most of them were already issues in the
%git bug reports. For example, we discovered that while the ternary
%operator is implemented in the compiler it cannot handle complex
%bounds types in the branches. The static analysis is not sophisticated
%enough to properly detect that both branches have the same type. While
%not precisely a bug, the clang compiler does not permit memory for
%null terminated arrays to be allocated with calloc. Although calloc
%fills all spaces in memory with null, the compiler does not recognize
%this and claims that it is an unsafe cast.

>>>>

% \review{While I found the idea behind section V very interesting, the current version
%   of this section lacks some details that would help in better understanding (1) 
%   how the approach works, and (2) the overall scope of the approach. 
%   $\\$
%   For instance, the authors state that, following [19], they try to "exercise
%   interesting patterns" by adding "admissible but redundant typing rules" like
%   G-ASTR. There are a few points that are unclear here: (1) are these rules
%   discovered manually or automatically (starting from the Redex semantics)?, (2)
%   are there any guiding principles for coming up with rules that lead to
%   interesting cases?
%   $\\$
%   Later, the authors refer to "generation rules modified to be slightly more
%   permissive" to generate "a little" ill-typed terms. Again, are these rules
%   obtained automatically or defined manually? If the latter, did you follow any
%   methodology to derive such rules? Are these rules the same as the "admissible
%   but redundant typing rules" from above?}
% \liyi{Deena? Leo? }
Our evaluation of \systemname consists of a series of tests that can be categorized into Micro-benchmarks and Program-benchmarks.
MicroBenchmarking involves evaluating performance on fundamental operations involving tainted pointers, context switching between checked and sandboxed regions, and sandboxed execution of functions. We further go on to evaluate \systemname on six real-world programs pertaining to diversified domains to evaluate real-world run-time and memory performance. We use C's \<time.h\> library to evaluate the Runtime performance by placing each of the test-bench calls within the timing scope (Program scope within which the timer runs). Our timing scope excludes marshaling activities within the test-case environment with an intuition that marshaling is irrelevant if tainted\-ness of a pointer is propagated from the call site up until its declaration. We use Valgrind's "massif" memory profiler to benchmark the peak memory usage of the Heap. Unlike the Runtime performance, we record Peak Memory usage as a relative offset to original program because most programs are extremely small in comparison to the constant overhead from the sandbox (81 KiB approx). This can further be optimized away by choosing to only compile custom Tainted wrappers for those STDLib functions that are in use by the tainted pointers in the program. Valgrind's memory figures do not account for Sandboxed allocations which happen in the shadow memory. 
All of the evaluation was performed using 6-Core Intel i7-10700H with 40 GB of RAM, running Ubuntu 20.04.3 LTS and the benchmarks for every test were sampled as the mean of ten consecutive iterations.

\subsection{Micro-Benchmarks}
\subsubsection{Memory Access in SBX}
Memory access performance between the checked and the WASM region involves crafting a test case that involves a simple pointer arithmetic operation enclosed in a loop of 100k iterations. 156.6\% overhead with \systemname is caused by the code executing in WASM Sandbox which is comparatively inefficient as WASM compiler toolchain does not support code optimization like LLVM/GCC.

\subsubsection{Indirect-Call}
This test involves evaluating the overhead involved in making indirect calls and is recorded by benchmarking the time taken for a sandboxed Callee (call from the checked region into the SBX) to return. This metric is evaluated against the time taken for a checked Callee (call within local scope) to return. Observed overhead is once again a consequence of code executing in the WASM Sandbox and call indirection.

\subsubsection{Pointer Access}
Pointer overhead during run-time between the pointers in the checked and Sandboxed regions is evaluated by benchmarking a test that involves 100k read/write/arithmetic operations on the pointer. 34\% in overhead is the result of "Offset To Pointer" conversion and sanity checks for Taintedness inserted by \systemname at every access to the tainted pointer.

\begin{center}
\begin{tabular}{||c c||} 
 \hline
 Test-Name & \systemname overhead \\ [0.5ex] 
 \hline\hline
 Memory Access in SBX & +156\% \\
 Indirect-Call & +398\% \\ 
 Pointer Access & +34.16\% \\ [1ex]
 \hline 
\end{tabular}
\end{center}

\subsection{Program Run-time Benchmarks}
We evaluate \systemname on the basis of conversion efforts and performance on six programs with the use of a pre-included test suite for each program. In our conversion efforts, we intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern-day memory vulnerabilities by re-introducing fixed memory bugs and then annotating the relevant buggy code with tainted pointers. However, this requires the bugs to be predicted before they are discovered, which is chronologically impossible. Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations without any input on the bug locality. The extent of conversion qualitatively dictates the converted program's conversion efforts required and the run-time performance. Consequently, each of the conversions for the six programs follows a varied approach. Parsons is annotated comprehensively in two variants parsons\_wasm and parsons\_tainted. parsons\_wasm has most of its input parsing functions moved into the sandbox, whilst having all its pointers marked as tainted. These sandboxed functions interact with the checked region by making indirect calls through RLBOX's callback mechanism. However, with parsons\_tainted, we do not move any of the functions to the sandbox but still mark all the pointers as tainted. The test suite itself consists of 328 tests comprehensively testing the JSON parser's functionality. Benchmarks for both of these forks are recorded using the mean difference between the \systemname and generic-C/checked-C variants when executing 10 consecutive iterations of the test suite. parsons\_wasm expectedly shows 200/266\% runtime overhead when evaluated against checked-c and generic-c respectively due to the performance limitation of WebAssembly. However, evaluating parsons\_tainted against checked-c shows \systemname to be faster because \systemname by itself performs lighter run-time-instrumentation on tainted pointers as compared to the run-time bounds checking performed on checked pointers by checked-c. Furthermore, we only see an average peak memory of 9.5 KiB as compared to the anticipated 82 KiB overhead as Valgrind does not consider the WASM Shadow memory allocated to the tainted pointers.

We intended to keep \systemname changes for Lib-png comparatively narrow in scope by beginning with the encapsulation CVE-2018-144550 and a buffer overflow in compare\_read(). However, we also annotate sections of Lib-png that involve reading, writing, and image processing (interlace, intrapixel, etc) on user-input image data as tainted. That is, rows of image bytes are read into tainted pointers and the taintedness for the row\_bytes is propagated throughout the program. All our changes extend to the png2pnm and pnm2png executables. To evaluate png2pnm, we take the mean of 10 iterations of a test script that runs png2pnm on 52 png files located within the libpng's pngsuite. To test pnm2png, we take the mean of 10 iterations of pnm2png in converting a 52MB 5184x3456 pixels large pnm image file to png. Valgrind's reported lower Heap space consumption for \systemname converted code is due to the discounted space consumed on the heap by the Sandbox's shadow memory. Consequently, when evaluating pnm2png, \systemname's heap consumption was 52 MB lower as the entire image was loaded onto the shadow memory.

\systemname changes for ProFTPD are further limited in extent and aimed at the exact changes required to encapsulate CVE-2010-4221. We mark the user input to the unsafe function "pr\_netio\_telnet\_gets()" as tainted (\_TPtr\<char\>) and propagate its tainted-ness to its callers and callees enclosed within its defined scope. Our changes for the above function were policed by 24 unique API tests, which we use to benchmark the performance. For each test, our benchmark samples the delta between the call and return time of pr\_netio\_telnet\_gets(). This sampling is repeated 10 times and its mean value is reflected in the below table. Although \systemname is shown to require 50\% more time in servicing this request, this delta is made insignificant during deployment where network bandwidth relatively accounts for a bigger metric to the overall performance.

MicroHTTPD is a good demonstration of the practical difficulties in converting a program to \systemname. Our conversion for this program was aimed at sandboxing memory vulnerabilities CVE-2021-3466 and CVE-2013-7039. CVE-2021-3466 is described as a vulnerability from a buffer overflow that occurs in an unguarded "memcpy" which copies data into a structure pointer (struct MHD\_PostProcessor pp) which is type-casted to a char buffer (char *kbuf = (char *) \&pp[1]). Our changes would require making the "memcpy" safe by marking this pointer as tainted. However, this would either require marshaling the data pointed by this structure (and its sub-structure pointer members) pointer or would require marking every reference to this structure pointer as tainted, which in turn requires every pointer member of this structure to be tainted. Marshalling data between structure pointers is not easy and demands substantial marshaling code due to the spatial non-linearity of its pointer members unlike a char*. This did not align with our conversion goals which were aimed at making minimal changes. Consequently, the above CVE stands un-handled by \systemname.  Our changes for CVE-2013-7039 involve marking the user input data arguments of this function as tainted pointers and in the interests of seeking minimal conversion changes, we do not propagate the tainted-ness on these functions. Following up on the chronological impossibility of sandboxing bugs before they are discovered and the general programmer intuiting, we moved many of the core internal functions (like MHD\_str\_pct\_decode\_strict\_() and MHD\_http\_unescape()) into the sandbox.

\systemname changes for UFTPD were aimed at sandboxing CVE-2020-14149 and CVE-2020-5204. CVE-2020-14149 was recorded as a NULL pointer dereference in the handle\_CWD() which could have led to a DoS in versions before 2.12, thereby, requiring us to sandbox this function. CVE-2020-5204 was recorded as a buffer overflow vulnerability in the handle\_PORT() due to sprintf() which also required us to sandbox this function. Although we could have chosen to only mark the faulty pointers as tainted, we intended to keep our changes more generic. For evaluation, we manually write a script for 3 Tests that each trigger "quote CWD", "quote PORT", and FTP "get file" request 10 times in a loop. Following this we record an entry for each of these tests as a mean of recorded timestamps for 10 executions for \systemname and generic-c versions, following which we record the relative average latency across the three tests between both of these UFTPD versions as a percentage in the below table.

Motivated by its size and simplicity, we intended to make comprehensive \systemname changes for Tiny-bignum by marking all the pointers as tainted. Furthermore, in likes of an unsafe use of sprintf() that could possibly lead to a buffer overflow within the bignum\_to\_string(), we move this function into the WASM sandbox. Ideally, a major portion of conversion efforts to \systemname is attributed to understanding the codebase and finding the precise extent to which we choose to propagate the taintedness or to stop and give up and marshall the data, thereby, requiring only 4 hours to convert tiny-bignum to \systemname. The evaluation was performed on Tiny-bignum's test suite consisting of 4 Test cases, each of which test the functionality to scale on big numbers subject to all of the supported unary and binary operations.

\begin{center}
\begin{tabular}{||c | p{1.3cm} | p{1.8cm}||} 
 \hline
 Program & Run-Time & Avg Peak Memory \\ [0.5ex] 
 \hline\hline
 ProFTPD & +47.8\% & +82.1 KiB  \\
 MicroHTTPD & +171.7\% & +24.7 KiB\\ 
 UFTPD & +2.7\% & +86.3 KiB \\ 
 Tiny-Bignum (vs checked-c) & -23.4\% &  +81.1 KiB \\
 Tiny-Bignum &  +121.7\% &  +81.1 KiB \\
 parsons\_wasm (vs checked-c) & +200.2\% & +9.8 KiB\\ 
 parsons\_tainted (vs checked-c) & -8.7\% & +9.2 KiB\\
 parsons\_wasm & +266.5\% & +9.8 KiB\\ 
 parsons\_tainted & +10.45\% & +9.2 KiB\\ 
 LibPNG (png2pnm) & +11.41\% & -102.5 KiB\\ 
 LibPNG (pnm2png) & +46.5\% & -51.6 MiB\\ [1ex]
 \hline 
\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{||p{2 cm} | p{1.2cm} | p{1.2cm} | p{1.0cm} | p{1.0cm}||} 
 \hline
 Program & Pointers Annotated  & Lines Sandboxed & CVEs fixed & Time To Port \\ [0.5ex] 
 \hline\hline
 ProFTPD & 6 & 0 & - & 1 Day  \\
 \hline
 MicroHTTPD & 139 & 650 & 1 & 4 Days \\
 \hline
 UFTPD & 146 & 90 & 2 & 3 Days \\ 
 \hline
 LibPNG & 248 &  0 & 1 & 8 Days \\
 \hline
 Tiny-Bignum &  69 &  30 & - & 4 Hours \\
 \hline
 parsons\_wasm & 364 & 800 &  - & 3 Days \\ 
 \hline
 parsons\_tainted & 378 & 0 & - & 8 Days \\ [1ex]
 \hline 
\end{tabular}
\end{center}
% 
%\leo{The following is extremely
%  weak. ``Most of them'', were there any that weren't? Which ones? For
%  the ones that were, mention github issues.}  The random generator,
%equipped with the conversion tool, successfully found a few minor
%errors in the clang compiler, most of them were already issues in the
%git bug reports. For example, we discovered that while the ternary
%operator is implemented in the compiler it cannot handle complex
%bounds types in the branches. The static analysis is not sophisticated
%enough to properly detect that both branches have the same type. While
%not precisely a bug, the clang compiler does not permit memory for
%null terminated arrays to be allocated with calloc. Although calloc
%fills all spaces in memory with null, the compiler does not recognize
%this and claims that it is an unsafe cast.
